Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-05-14 _tex/paper2.tex
Built site for gh-pages

Next, we specify the probabilistic assumptions for the idiosyncratic
errors \(e_{I}\), \(e_{IA}\), \(e_{J}\), and \(e_{JK}\). These error
terms exhibit indeterminacies in their \emph{location},
\emph{orientation}, and \emph{scale} due to the lack of an inherent
scale in the associated latent variables. We must resolve these
indeterminacies to identify the model
\citep{Depaoli_2021, deAyala_2009}. Drawing on principles from SEM
\citep{Hoyle_et_al_2023}, we assume that the vector of idiosyncratic
errors \(\boldsymbol{e} = [e_{I}, e_{IA}, e_{J}, e_{JK}]^{T}\), follows
a Multivariate Normal distribution with mean vector \(\boldsymbol{\mu}\)
and a covariance matrix
\(\boldsymbol{\Sigma} = \boldsymbol{D} \boldsymbol{R} \boldsymbol{D}\),
with \(\boldsymbol{D}\) denoting a diagonal matrix of standard
deviations and \(\boldsymbol{R}\) a correlation matrix. To address the
\emph{location} indeterminacy, we set the errors' mean vector to zero:
\begin{equation}\phantomsection\label{eq-mat53}{
\boldsymbol{\mu} = [0, 0, 0, 0]^{T}
}\end{equation}

Following SCM~\ref{fig-cj16_scm}, we assume uncorrelated errors to
resolve the \emph{orientation} indeterminacy. Thus, we define the
correlation matrix \(\boldsymbol{R}\) as:
\begin{equation}\phantomsection\label{eq-mat54}{
\boldsymbol{R} = \begin{bmatrix}
    1 & 0 & 0 & 0 \\
    0 & 1 & 0 & 0 \\
    0 & 0 & 1 & 0 \\
    0 & 0 & 0 & 1 
  \end{bmatrix}
}\end{equation}

@@

Here, \(s_{XI}\) represents the standard deviation for the individuals,
\(p_{IA}\) for the stimuli, \(s_{ZJ}\) for the judges, and \(p_{JK}\)
for the judgments. We assume \(s_{XI}\) varies depending on the teaching
method group to which each student belongs. In the example from
Section~\ref{sec-theoretical}, where the teaching method
\(X_{I} = \{1,2\}\), the model sets the constraint according to Equation
\ref{eq-mat56}. This constraint anchors the scale of the individuals'
latent trait while relaxing the assumption of equal dispersion for the
stimuli, thereby addressing the concerns raised in
Section~\ref{sec-theory-issue1a}.
\begin{equation}\phantomsection\label{eq-mat56}{
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
}\end{equation}

@@

Similarly, we assume \(s_{ZJ}\) varies depending on the groups to which
each judge belongs. For instance, if \(Z_{J} = \{1,2,3\}\) represents
three groups of judges with varying expertise, the model sets the
constraint: \begin{equation}\phantomsection\label{eq-mat56}{
\sum_{g=1}^{2} s_{ZJ}[g]/3 = 1
}\end{equation}

This constraint anchors the scale of the judges' latent trait and
relaxes the assumption of equal dispersion for the judgments.
Conversely, \(p_{JK}\) is defined as a proportion of \(1\) to establish
the scale of the judgments' latent trait relative to the scale of the
judges' trait.

>>>>

Next, we specify the probabilistic assumptions for the idiosyncratic
errors \(e_{I}\), \(e_{IA}\), \(e_{J}\), and \(e_{JK}\). These error
terms exhibit indeterminacies in their \emph{location},
\emph{orientation}, and \emph{scale} due to the lack of an inherent
scale in the associated latent variables. Thus, we must resolve these
indeterminacies to identify the latent variable model
\citep{Depaoli_2021, deAyala_2009}. Drawing on principles from SEM
\citep{Hoyle_et_al_2023}, we assume that the vector of idiosyncratic
errors \(\boldsymbol{e} = [e_{I}, e_{IA}, e_{J}, e_{JK}]^{T}\), follows
a Multivariate Normal distribution with mean vector \(\boldsymbol{\mu}\)
and a covariance matrix
\(\boldsymbol{\Sigma} = \boldsymbol{D} \boldsymbol{R} \boldsymbol{D}\),
with \(\boldsymbol{D}\) denoting a diagonal matrix of standard
deviations and \(\boldsymbol{R}\) a correlation matrix. To address the
\emph{location} indeterminacy, we set the errors' mean vector to zero:
\begin{equation}\phantomsection\label{eq-mat53}{
\boldsymbol{\mu} = [0, 0, 0, 0]^{T}
}\end{equation}

Following SCM~\ref{fig-cj16_scm}, we resolve the \emph{orientation}
indeterminacy by assuming that the errors are uncorrelated. This
assumption leads us to specify a correlation matrix for the errors,
\(\boldsymbol{R}\), as: \begin{equation}\phantomsection\label{eq-mat54}{
\boldsymbol{R} = \begin{bmatrix}
    1 & 0 & 0 & 0 \\
    0 & 1 & 0 & 0 \\
    0 & 0 & 1 & 0 \\
    0 & 0 & 0 & 1 
  \end{bmatrix}
}\end{equation}

@@

Here, \(s_{XI}\) represents the standard deviation for the individuals,
\(p_{IA}\) for the stimuli, \(s_{ZJ}\) for the judges, and \(p_{JK}\)
for the judgments. We assume \(s_{XI}\) varies depending on the teaching
method group to which each student belongs. In the example from
Section~\ref{sec-theoretical}, where the teaching method
\(X_{I} = \{1,2\}\), the model sets the constraint according to Equation
(\ref{eq-mat56}). This constraint anchors the scale of the individuals'
latent trait while relaxing the assumption of equal dispersion for the
stimuli, thereby addressing the concerns raised in
Section~\ref{sec-theory-issue1a}.
\begin{equation}\phantomsection\label{eq-mat56}{
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
}\end{equation}

@@

Similarly, we assume \(s_{ZJ}\) varies depending on the groups to which
each judge belongs. For instance, if \(Z_{J} = \{1,2,3\}\) represents
three groups of judges with varying expertise, the model sets the
constraint according to Equation (\ref{eq-mat57}). This constraint
anchors the scale of the judges' latent trait and relaxes the assumption
of equal dispersion for the judgments.
\begin{equation}\phantomsection\label{eq-mat57}{
\sum_{g=1}^{2} s_{ZJ}[g]/3 = 1
}\end{equation}

Conversely, \(p_{JK}\) is defined as a proportion of \(1\) to establish
the scale of the judgments' latent trait relative to the scale of the
judges' trait.
